On the power of oritatami cotranscriptional folding with unary bead sequence\thanks{This work is supported in part by KAKENHI Grant-in-Aid for Challenging Research (Exploratory) No.~18K19779 and JST Program to Disseminate Tenure Tracking System No.~6F36 granted to S.~Z.~F. and S.~S.}

### 2018-12-06 paper.tex
space was missing
Time hierarchy

\end{tikzpicture}
\caption{Upper bounds on the length of a conformation foldable by a deterministic unary oritatami system at delay $\delta$ and arity $\alpha$.
At any combination of delay and arity without anything written, no upper bound is known yet.
}
\label{fig:summary}
\end{figure}
In this paper, we will solve this problem  completely at delay 1and partially at arity 1 in Sect.~\ref{sect:upper}. 
At delay 1, we will provide a quadratic upper bound $3n(n+1)+1$ for the case of arity being 4, while a linear upper bound $4n+14$ for arity 3. 
At the delay 1 and arity 2, one infinite structure turns out to be foldable deterministically, which is the zigzag conformation shown in Fig.~\ref{TTT_zigzag}. 
At arity 1, we will prove that at delay 2 or 3, the upper bound is $(3n+1)(n+1)$. 
Upper bounds for longer delays remain open. These results as well as known upper bounds are summarized in Fig.~\ref{fig:summary}. 
As shown at the end of Sect.~\ref{sec:preliminaries}, the stabilization of first $t$ beads of transcript by a determinisitc oritatami system can be simulated by a deterministic Turing machine within $t^3$ steps (Corollary~\ref{cor:DTIME3}). 
Thus, the above mentioned upper bounds show that at delay 1 and arity 1, 3, or 4, or at delay 2 or 3 and arity 1, the class of deterministic unary oritatami systems is not Turing universal. 
The Turing universal oritatami system by Geary et al. \cite{GeMeScSe2018} employs more than 500 types of beads. 
Thus, this weakness result is not surprising at all. 
The unary oritatami system might not be practical very much, though one bead may abstarct an oligonucleotide (a short sequence of nucleotides), and in that case, unary transcript can be a repetitive but nonunary sequence, which is not so unrealistic in experiments (see \cite{GearyAndersen2014}). 
Nevertheless, this paper makes a first considerable step towards the characterization of non-Turing-universal oritatami systems.

@@

Oritatami systems fold their transcript, which is a sequence of beads, over the triangular grid graph $\mathbb{T} = (V, E)$ cotranscriptionally. 
We designate one point in $V$ as the origin $O$ of $\mathbb{T}$. 
For a point $p \in V$, let $\hexagon_p^d$ denote the set of points which lie in the regular hexagon of radius $d$ centered at the point $p$. 
Note that $\hexagon_p^d$ consists of $3d(d+1)+1$ points. 
A directed path $P = p_1 p_2 \cdots p_n$ in $\mathbb{T}$ is a sequence of \textit{pairwise-distinct} points $p_1, p_2, \ldots, p_n \in V$ such that $\{p_i, p_{i+1}\} \in E$ for all $1 \le i < n$. 
Its $i$-th point is referred to as $P[i]$. 
Now we are ready to abstract RNA single-stranded structures in the name of conformation. 
A \textit{conformation} $C$ (over $\Sigma$) is a triple $(P, w, H)$ of a directed path $P$ in $\mathbb{T}$, $w \in \Sigma^*$ of the same length as $P$, and a set of h-interactions $H \subseteq \bigl\{\{i, j\} \bigm| 1 \le i, i+2 \le j, \{P[i], P[j]\} \in E \bigr\}$. 
This is to be interpreted as the sequence $w$ being folded along the path $P$ in such a manner that its $i$-th bead $w[i]$ is placed at the $i$-th point $P[i]$ and the $i$-th and $j$-th beads are bound (by a hydrogen-bond-based interaction) if and only if $\{i, j\} \in H$. 
The condition $i+2 \le j$ represents the topological restriction that two consecutive beads along the path cannot be bound. 
The \textit{length} of $C$ is defined to be the length of its transcript $w$ (that is, equal to the length of the path $P$). 
A \textit{rule set} $R \subseteq \Sigma \times \Sigma$ is a symmetric relation over $\Sigma$, that is, for all bead types $a, b \in \Sigma$, $(a, b) \in R$ implies $(b, a) \in R$. 
A bond $\{i, j\} \in H$ is \textit{valid with respect to $R$}, or simply $R$-valid, if $(w[i], w[j]) \in R$. 
This conformation $C$ is \textit{$R$-valid} if all of its bonds are $R$-valid. 
For an integer $\alpha \ge 1$, $C$ is \textit{of arity $\alpha$} if it contains a bead that forms $\alpha$ bonds but none of its bead forms more. 
By $\mathcal{C}_{\le \alpha}(\Sigma)$, we denote the set of all conformations over $\Sigma$ whose arity is at most $\alpha$; its argument $\Sigma$ is omitted whenever $\Sigma$ is clear from the context.

@@

Because of the time hierarchy theorems, we know that $\mathrm{P}\subsetneq \mathrm{EXP}$, so we can conclude that OS which cannot deterministically fold transcripts of length exponential in the length of the seed are not computationally universal.

>>>>

\end{tikzpicture}
\caption{Upper bounds on the length of a conformation foldable by a deterministic unary oritatami system at delay $\delta$ and arity $\alpha$.
At any combination of delay and arity without anything written, no upper bound is known yet.
}
\label{fig:summary}
\end{figure}
In this paper, we will solve this problem  completely at delay 1 and partially at arity 1 in Sect.~\ref{sect:upper}. 
At delay 1, we will provide a quadratic upper bound $3n(n+1)+1$ for the case of arity being 4, while a linear upper bound $4n+14$ for arity 3. 
At the delay 1 and arity 2, one infinite structure turns out to be foldable deterministically, which is the zigzag conformation shown in Fig.~\ref{TTT_zigzag}. 
At arity 1, we will prove that at delay 2 or 3, the upper bound is $(3n+1)(n+1)$. 
Upper bounds for longer delays remain open. These results as well as known upper bounds are summarized in Fig.~\ref{fig:summary}. 
As shown at the end of Sect.~\ref{sec:preliminaries}, the stabilization of the first $t$ beads of transcript by a deterministic oritatami system can be simulated by a deterministic Turing machine within $t^3$ steps (Corollary~\ref{cor:DTIME3}). 
Thus, the above mentioned upper bounds show that at delay 1 and arity 1, 3, or 4, or at delay 2 or 3 and arity 1, the class of deterministic unary oritatami systems is not Turing universal. 
The Turing universal oritatami system by Geary et al. \cite{GeMeScSe2018} employs more than 500 types of beads. 
Thus, this weakness result is not surprising at all. 
The unary oritatami system might not be practical very much, though one bead may abstract an oligonucleotide (a short sequence of nucleotides), and in that case, unary transcript can be a repetitive but nonunary sequence, which is not so unrealistic in experiments (see \cite{GearyAndersen2014}). 
Nevertheless, this paper makes a first considerable step towards the characterization of non-Turing-universal oritatami systems.

@@

Oritatami systems fold their transcript, which is a sequence of beads, over the triangular grid graph $\mathbb{T} = (V, E)$ cotranscriptionally. 
We designate one point in $V$ as the origin $O$ of $\mathbb{T}$. 
For a point $p \in V$, let $\hexagon_p^d$ denote the set of points which lie in the regular hexagon of radius $d$ centered at the point $p$. 
Note that $\hexagon_p^d$ consists of $3d(d+1)+1$ points. 
A directed path $P = p_1 p_2 \cdots p_n$ in $\mathbb{T}$ is a sequence of \textit{pairwise-distinct} points $p_1, p_2, \ldots, p_n \in V$ such that $\{p_i, p_{i+1}\} \in E$ for all $1 \le i < n$. 
Its $i$-th point is referred to as $P[i]$. 
Now we are ready to abstract RNA single-stranded structures in the name of conformation. 
A \textit{conformation} $C$ (over $\Sigma$) is a triple $(P, w, H)$ of a directed path $P$ in $\mathbb{T}$, $w \in \Sigma^*$ of the same length as $P$, and a set of h-interactions $H \subseteq \bigl\{\{i, j\} \bigm| 1 \le i, i+2 \le j, \{P[i], P[j]\} \in E \bigr\}$. 
This is to be interpreted as the sequence $w$ being folded along the path $P$ in such a manner that its $i$-th bead $w[i]$ is placed at the $i$-th point $P[i]$ and the $i$-th and $j$-th beads are bound (by a hydrogen-bond-based interaction) if and only if $\{i, j\} \in H$. 
The condition $i+2 \le j$ represents the topological restriction that two consecutive beads along the path cannot be bound. 
The \textit{length} of $C$ is defined to be the length of its transcript $w$ (that is, equal to the length of the path $P$). 
A \textit{rule set} $R \subseteq \Sigma \times \Sigma$ is a symmetric relation over $\Sigma$, that is, for all bead types $a, b \in \Sigma$, $(a, b) \in R$ implies $(b, a) \in R$. 
A bond $\{i, j\} \in H$ is \textit{valid with respect to $R$}, or simply $R$-valid, if $(w[i], w[j]) \in R$. 
This conformation $C$ is \textit{$R$-valid} if all of its bonds are $R$-valid. 
For an integer $\alpha \ge 1$, $C$ is \textit{of arity $\alpha$} if it contains a bead that forms $\alpha$ bonds but none of its beads forms more. 
By $\mathcal{C}_{\le \alpha}(\Sigma)$, we denote the set of all conformations over $\Sigma$ whose arity is at most $\alpha$; its argument $\Sigma$ is omitted whenever $\Sigma$ is clear from the context.

@@

Considering the following decision problem: given an oritatami system, integer $i$, and a point $p$, decides whether the bead $w[i]$ is stabilized at $p$. 
By Corollary~\ref{cor:DTIME3}, this problem is in $\mathrm{P}$ for a fixed delay $\delta$.
Because of the time hierarchy theorems, we know that $\mathrm{P}\subsetneq \mathrm{EXP}$ (see, e.g., \cite{AroraBarak2009}), so we can conclude that OS which cannot deterministically fold transcripts of length exponential in the length of the seed are not computationally universal.
